---
title: "Residential Regression R Code"
author: "Brendan Callender, Martin Hsu, and Rachel Roggenkemper"
date: "5/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(leaps)
library(car)
library(scatterplot3d)
library(GGally)
library('lmtest')
```

```{r}
realestate <- read_excel("SLO_Real_Estate_Feb-Apr2022.xlsx")

realestate_clean <- realestate %>%
  filter(LotSize != 416869 & LotSize != 108900 & LotSize != 84071)

realestate_clean <- realestate_clean %>%
  mutate(HomeTypeNew = 
           case_when(HomeType == "Condominium" ~ "Condo", 
             HomeType == "SingleFamilyResidence" ~ "SingleFam",
             HomeType == "ManufacturedHome" ~ "Manufactured",
             TRUE ~ "Other")) 

realestate_clean$Cooling = factor(realestate_clean$Cooling)
realestate_clean$Heating = factor(realestate_clean$Heating)
realestate_clean$HomeType = factor(realestate_clean$HomeType)
realestate_clean$HomeTypeNew = factor(realestate_clean$HomeTypeNew)
realestate_clean$Pool = factor(realestate_clean$Pool)
```



# IV. Split the Data
```{r}
n <- 141
numtrain = ceiling(.8*n)
set.seed(678)
train_ind = sample(n, numtrain)
traindata = realestate_clean[train_ind, ]
testdata = realestate_clean[-train_ind, ]
set.seed(NULL)
```



# V. Data Visualization
```{r}
# Matrix Scatterplot (outliers not removed)
pairs(SoldPrice ~ Sqft + Bed + Bath + YearBuilt + LotSize + Parking, data = realestate, 
      col = c('cornflowerblue','royalblue','cyan', 'darkseagreen')[factor(realestate_clean$HomeTypeNew)], 
      pch = 19)
```

```{r}
# Correlation Matrix (outliers not removed)
cor(realestate[,c(1, 2, 3, 4, 5, 6, 7)])
```

```{r}
# Matrix Scatterplot with ggpairs()
ggpairs(traindata[,-c(8:12)], 
        upper=list(continuous='points'),
        lower=list(continuous='cor',combo='blank',discrete='blank'))  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

```{r}
# Interaction Plot Between Sold Price and Square Footage (Home Type)
fit_interaction = lm(SoldPrice~Sqft*HomeTypeNew, data=traindata)
summary(fit_interaction)

plot(SoldPrice ~ Sqft, data = traindata, 
     col = c('cornflowerblue','royalblue','cyan', 'darkseagreen')[HomeTypeNew], 
     pch = c(15, 16, 17, 18)[HomeTypeNew])
legend('bottomright',legend=levels(traindata$HomeTypeNew),
       pch = c(15, 16, 17, 18),
       col = c('cornflowerblue','royalblue','cyan', 'darkseagreen'),
       lty = 1)
abline(coeff[1], coeff[2], col='cornflowerblue')
abline(coeff[1] + coeff[3], coeff[2] + coeff[6], col = 'royalblue') 
abline(coeff[1] + coeff[4], coeff[2] + coeff[7], col = 'cyan') 
abline(coeff[1] + coeff[5], coeff[2] + coeff[8], col = 'seagreen') 
```

```{r}
# Histograms of Quantitative Variables 
traindata %>%
  ggplot(aes(x = Sqft)) +
  geom_histogram(bins = 10) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

traindata %>%
  ggplot(aes(x = Sqft, fill = HomeTypeNew)) +
  geom_histogram(bins = 10) +
  scale_fill_manual(values = c('aquamarine', 'cornflowerblue','royalblue', 'cadetblue', 'cyan', 
                               'darkblue', 'deepskyblue', 'darkseagreen', 'darkseagreen1')) +
  facet_wrap(~HomeTypeNew) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

traindata %>%
  ggplot(aes(x = LotSize)) +
  geom_histogram(binwidth = 2000) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

traindata %>%
  ggplot(aes(x = LotSize, fill = HomeTypeNew)) +
  geom_histogram(binwidth = 2000) +
  scale_fill_manual(values = c('aquamarine', 'cornflowerblue','royalblue', 'cadetblue', 'cyan', 
                               'darkblue', 'deepskyblue', 'darkseagreen', 'darkseagreen1')) +
  facet_wrap(~HomeTypeNew) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

```{r}
# Interaction Plot Between Home Type and Pool
with(traindata, interaction.plot(HomeTypeNew, Pool, SoldPrice, type = 'b', 
                                 col = c('royalblue','cyan'), 
                                 pch = c(16, 17)))
with(traindata, interaction.plot(Pool, HomeTypeNew, SoldPrice, type = 'b', 
                                 col = c('cornflowerblue','royalblue','cyan', 'darkseagreen'), 
                                 pch = c(16, 17)))
```

```{r}
# 3D Scatterplot 
s3d <- scatterplot3d(traindata2$Sqft, y=traindata2$SoldPrice, z=traindata2$LotSize,
                     xlab = "Square Footage (sqft)",
                     ylab = "Sold Price (Dollars)",
                     zlab = "Lot Size (sqft)", 
                     angle=55, pch = 16)
# Add regression plane
my.lm <- lm(traindata2$Sqft ~ traindata2$SoldPrice + traindata2$LotSize)
s3d$plane3d(my.lm)
```


# Part VI. Variable Pre-Processing
Using Best Subsets, a model containing Sqft, YearBuilt, and HomeTypeNew looked like the best option for the project.
```{r}
traindataNew <- traindata %>%
  select(-HomeType)

full.model <- lm(SoldPrice~., data = traindataNew)

best.sub <- summary(regsubsets(formula(full.model), data=traindata, method = "exhaustive", nbest = 2))
```

```{r}
n = nrow(traindata)
p<-c(2,2,3,3,4,4,5,5,6,6,7,7,8,8,9,9) 

SSE<-best.sub$rss             #SSE 
s<-sqrt(SSE/(n-p))            #s
Rsq<-best.sub$rsq             #R-squared 
Rsq.adj<-best.sub$adjr2       #R-squared adjusted
BIC<-n*log(SSE/n) + p*log(n)  #BIC
AIC<- n*log(SSE/n) + 2*p      #AIC
Cp<-best.sub$cp               #Mallow's Cp
bs<- data.frame(best.sub$outmat, p=p, Cp, AIC, BIC, Rsq, Rsq.adj, s) %>%
  arrange(BIC)
```

Many models reccomended the inclusion of lot size however one AVPlot ruled out LotSize immediately.  You can see one point that is heavily influencing the entire trend.  This is very simular to one plot from Anscomb's Quartet. So the next best Model that met the project requirements 

```{r}
plot(resid(lm(I(sqrt(SoldPrice))~Sqft + HomeTypeNew, data = traindata))~resid(lm(LotSize ~ YearBuilt + Sqft + HomeTypeNew, data = traindata)))
```

```{r}
fit.one <- lm(SoldPrice ~ Sqft + LotSize + HomeTypeNew, data  = traindata)

plot(fit.one)
```

Looking at residual vs predicted plot, there is a obvious violation of the equal variance assumption as well as normality assumption.  We will likely need a transform the response and then transform a predictor if necessary.

Since we are transforming the y variable we can use the BoxCox procedure to see which transformation of the response minimizes the SSE.

```{r}
boxCox(lm(SoldPrice~Sqft+LotSize+HomeTypeNew, data = traindata))
```

Using Box-Cox, we decided a sqrt (1/2) transformation on the response variable of SoldPrice because it is contained in the confidence interval for the box cox transformation.

```{r}
fit.two <- lm(I(sqrt(SoldPrice)) ~ Sqft + LotSize + HomeTypeNew, data  = traindata)

plot(fit.two)
```

```{r}
full.model2 <- lm(I(sqrt(SoldPrice))~., data = traindataNew)

best.sub2 <- summary(regsubsets(formula(full.model2), data=traindata, method = "exhaustive", nbest = 2))
```

```{r}
n = nrow(traindata)
p<-c(2,2,3,3,4,4,5,5,6,6,7,7,8,8,9,9) 

SSE<-best.sub2$rss             #SSE 
s<-sqrt(SSE/(n-p))             #s
Rsq<-best.sub2$rsq             #R-squared 
Rsq.adj<-best.sub2$adjr2       #R-squared adjusted
BIC<-n*log(SSE/n) + p*log(n)   #BIC
AIC<- n*log(SSE/n) + 2*p       #AIC
Cp<-best.sub2$cp               #Mallow's Cp
bs2<- data.frame(best.sub2$outmat, p=p, Cp, AIC, BIC, Rsq, Rsq.adj, s) %>%
  arrange(BIC)
```


# VII. Residual Analysis
```{r}
final_model = lm(I(sqrt(SoldPrice)) ~ Sqft + LotSize + HomeTypeNew, data = traindata)

par(mfrow=c(2,2)) #plot in 2 by 2 grid   
plot(final_model)

par(mfrow=c(2,2)) #plot in 2 by 2 grid  

plot(resid(final_model) ~ fitted(final_model), ylab='Residuals', xlab='Fitted')
abline(h=0,lty=2)

plot(resid(final_model) ~ Sqft, data=traindata, ylab='Residuals', xlab='Sqft')
abline(h=0,lty=2) 

plot(resid(final_model) ~ LotSize, data=traindata, ylab='Residuals', xlab='Lot Size')
abline(h=0,lty=2) 

plot(resid(final_model) ~ HomeTypeNew, data=traindata, ylab='Residuals', xlab='HomeTypeNew')
abline(h=0,lty=2)

par(mfrow=c(1,1))

hist(resid(final_model), main='', xlab='Residuals')

shapiro.test(resid(final_model))
bptest(final_model)
```



# VIII. Fit a Linear Model 
```{r}
summary(final_model)
vif(final_model)
```



# IX. Statistical Inference
```{r}
# Overall F-Test
summary(final_model)
```

```{r}
# Partial F-Test
model_reduced = lm(I(sqrt(SoldPrice)) ~ Sqft, data=traindata)
anova(model_reduced, final_model)
```

```{r}
# Test for Interaction Significance 
anova(lm(SoldPrice ~ Sqft + HomeTypeNew, data = traindata), 
      lm(SoldPrice ~ Sqft + HomeTypeNew + Sqft:HomeTypeNew, data = traindata))

anova(lm(I(sqrt(SoldPrice)) ~ LotSize + Sqft + HomeTypeNew, data = traindata), 
      lm(I(sqrt(SoldPrice)) ~ LotSize + Sqft + HomeTypeNew + Sqft:HomeTypeNew, data = traindata))

anova(lm(I(sqrt(SoldPrice)) ~ LotSize + Sqft + HomeTypeNew, data = traindata), 
      lm(I(sqrt(SoldPrice)) ~ LotSize + Sqft + HomeTypeNew + Sqft:LotSize, data = traindata))
```

```{r}
# CIs for Mean Response Value and Future Predicted Value 
new.house = data.frame(Sqft = 3000, LotSize = 10000, HomeTypeNew = "SingleFam")

predict(final_model, newdata = new.house, interval = "confidence", level=0.95)
predict(final_model, newdata = new.house, interval = "prediction", level=0.95)
```


